Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/intro.tex

Second,  applications often need to produce output in the same
format as their input---for example, applications might both
read and write files, or both receive and send network packets.
Most parser generators just focus on parsing an input, rather
than producing an output, thus requiring the programmer to manually
construct outputs, which is error-prone.  Some parser generators, such as
Boost.Spirit~\cite{boost-spirit},
allow reusing the grammar for generating output from the internal
representation.  However, those generators require yet another set of
semantic actions to be written, transforming the internal representation
into an AST\@.

Third, many data formats contain redundancies, such as repeating information in multiple structures.
Applications usually do not explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft input that causes inconsistencies.
Furthermore, security vulnerability can occur when an application assumes two repetitions of the
same data to be consistent, such as allocating a buffer based on one size field and copying into
that buffer based on another.

@@

This paper presents the design and implementation of Nail, a parser generator that greatly reduces
the programmer effort required to use grammars. Nail addresses the above four challenges with
several key ideas.

@@

Second, this well-defined internal representation allows Nail to generate the external format from
the internal representation, without requiring the programmer to write additional code.

@@

We have implemented a prototype of Nail for C\@.  Our experience
so far suggests that it is a promising approach: we were able to construct
a succinct grammar for a set of common file formats and protocols that demonstrate Nail's ability to
handle tricky constructs and build a series of sample applications with these grammars.

>>>>

Second,  applications often need to produce output in the same
format as their input---for example, applications might both
read and write files, or both receive and send network packets.
Most parser generators focus on just parsing an input, rather
than producing an output, thus requiring the programmer to manually
construct outputs, which is error-prone.  Some parser generators, such as
Boost.Spirit~\cite{boost-spirit},
allow reusing the grammar for generating output from the internal
representation.  However, those generators require yet another set of
semantic actions to be written, transforming the internal representation
into an AST\@.

Third, many data formats contain redundancies, such as repeating information in multiple structures.
Applications usually do not explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft input that causes inconsistencies.
Furthermore, security vulnerabilities can occur when an application assumes two repetitions of the
same data to be consistent, such as allocating a buffer based on the value of one size field and copying into
that buffer based on the value of another.

@@

This paper presents the design and implementation of Nail, a parser generator that greatly reduces
the programmer effort required to use grammars. Nail addresses the above four challenges with
several key ideas, as follows.

@@

Second, this well-defined internal representation allows Nail to
maintain a {\em semantic bijection} between data formats and their
internal representation.  As a result, this enables Nail to not just
parse input but also generate output from the internal representation,
without requiring the programmer to write additional code.

@@

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for Zip files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to handle external data formats.
